---
title: "Exercise 3"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
## clear memory
rm(list = ls())
gc()

## load the necessary packages
library(tidyverse)
library(skimr)
library(lubridate)
library(arrow)
set.seed(10101)

##### paths -----
data_path = "~/Library/CloudStorage/Dropbox/McGill/teaching/2024-2025/winter/talent-analytics/data/" 
```

## Load the data

Let's start by loading the project data file and looking at its structure.

```{r}
## load the data
app_data_raw <- read_feather(paste0(data_path,"app_data_starter_coded.feather")) 

## show the top 10 rows
app_data_raw
```

## Prepare the monthly aggregated panel

We will now aggregate the data on the monthly basis. Specifically, we need to follow these steps:

### First, examiner-level features

Start with creating a copy of the raw data to calculate these:

1.  Extract `year` and `month` from the `app_last_action` field
2.  Group data by `year`, `month` and `examiner_id`
3.  Summarize by the groups, to create the panel data, which will have these variables:
    -   total number of applications processed
    -   examiner tenure in days since hire date (approximated with `examiner_earliest_date`)
    -   examiner's art unit number in that month
    -   total number of applications with patents issued (`app_disposal_type == "ISS"`)
    -   total number of applications that were abandoned after a rejection (`app_disposal_type == "ABN"`)
    -   total number of applications with final decision still pending (`app_disposal_type == "PEND"`)
    -   also keep gender and race variables, though they don't change within examiner

```{r}
## your code here
```

### Calculate art-unit-level features

Create another copy of the raw data, but aggregate it on a different level

1.  Extract year and month fields again,
2.  Group data by year, month, and art unit number
3.  Summarize to create these variables
    -   number of examiners in that art unit in that month
    -   number of minority-race (non-white) examiners
    -   number of women
    -   number of applications processed in that month

```{r}
## your code here
```

### Join the art unit characteristics to the examiner monthly data

Join the second generated dataset to the first one using `left_join()` command in `dplyr` package, and calculate additional variables:

-   extract TC number from `au` into a `tc` variable. Remember that it's indicated by the first two digits, so an art unit 1652 is located in Technology Center 1600

-   extract workgroup (WG) number from art unit, using the first 3 digits of the art unit. So, an art unit 1652 is part of the workgroup `wg` 1650

```{r}
## your code here
```

## Estimate, using linear regressions, what factors predict productivity

Using the resulting monthly dataset, estimate what factors in the data are associated with higher/lower number of applications processed by any given examiner in any given month.

```{r}
## your code here
```
